---# Class 1 Survey Data Analysis
## Project Description
This repository includes the Class 1 survey dataset and R code used to
analyze it. The project is part of an exercise to practice working with real-
world data in R.
## Files Included
- `class1_survey.csv`: Cleaned version of the Class 1 survey dataset
- `analysis_code.R`: R script used to summarize and visualize the data
- `README.md`: This file
## What the Code Does
- Reads in the survey dataset
- Performs basic data cleaning (e.g., renaming columns, filtering)
- Creates summary statistics (e.g., age, gender, responses)
- Generates plots to visualize the data
## How to Run the Code
1. Download or clone this repository to your computer
2. Open `analysis_code.R` in RStudio
3. Make sure your working directory is set to the folder where the files are
saved
4. Run the script
## Author
- Name: Bezaye
- Course: Advanced data analysis
- Date: 9/7/2025

```{r, chunk1, message=FALSE}
#Import class one survey data from our Github site
#Save it as C1survey
library(readr)
C1survey_csv <-
 read_csv(
   "https://github.com/kijohnson/Advanced-Data-Analysis/raw/refs/heads/main/Class%201%20Survey%20Fall%202025%202.csv"
   )
```
```{r, chunk2}
# a. Number of observations (people who filled out the survey)
# b. Number of variables in the dataframe
n_obs <- nrow(C1survey_csv)
n_vars <- ncol(C1survey_csv)
```


```{r, chunk3}
#using the dimension function to find the number of obseravations
dim(C1survey_csv) 
```


```{r, chunk4}
View(C1survey_csv)
```
```{r}
install.packages("dplyr")
library(dplyr) 
```

```{r, chunk5}
# Load the dplyr function
library(dplyr) 
colnames(C1survey_csv) 
```
```{r, Chunk6}
#3. a. Rename the column variables to something shorter and descriptive
C1survey_csv <- C1survey_csv %>%
  rename(
    id = id,
    like_cats = like_cats,
    like_dogs = like_dogs,
    have_desert = have_desert,
    slogan = slogan,
    fav_day = fav_day,
    larkORowl = larkORowl,
    fav_food = fav_food,
    fav_drink = fav_drink,
    fav_season = fav_season,
    fav_month = fav_month,
    hobby = hobby,
    program = program,
    specialization = specialization,
    stat_software = stat_software,
    R_exp = R_exp,
    coding_comfort = coding_comfort,
    coding_length = coding_length,
    top_three = top_three,
    public_health_interest = public_health_interest,
    fav_num = fav_num,
    bday = bday,
    bmonth = bmonth,
    country = country,
    state = state,
    city = city,
    highest_educ_level = highest_educ_level
  )



```

```{r, chunk7}
# 3. b. Display new column names
colnames(C1survey_csv)
```
```{r, chunk8}
# 4. display the number of factor, integer, numerical, and character variables there are in the C1survey dataset. 
sapply(C1survey_csv, class) |> 
  table()
```
```{r, chunk9}
# 5. a. Check the bday and bmonth variables for unusual or missing values
table(C1survey_csv$bday, useNA = "always")
table(C1survey_csv$bmonth, useNA = "always")
```
```{r, chunk10}
#check the class
class(C1survey_csv$bday)
class(C1survey_csv$bmonth)
```
```{r, chunk11}
#check the class
class(C1survey_csv$bday)
class(C1survey_csv$bmonth)
```
```{r, chunk12}
#5. b. If there are invalid values (e.g. day < 1 or > 31, month < 1 or > 12), replace them with NA
C1survey_csv$bday <- as.numeric(C1survey_csv$bday)
C1survey_csv$bmonth <- as.numeric(C1survey_csv$bmonth)
```

```{r, chunk13}
#Clean NA for bday
C1survey_csv$bday[C1survey_csv$bday < 1 | C1survey_csv$bday > 31] <- NA

```
```{r, chunk14}
#Clean NA for bmonth
C1survey_csv$bmonth[C1survey_csv$bmonth < 1 | C1survey_csv$bmonth > 12] <- NA
```

```{r, chunk15}
#see the table after I Cleaned NA
table(C1survey_csv$bday, useNA = "always")
table(C1survey_csv$bmonth, useNA = "always")
```
```{r, chunk16}
#5. c. Find median birthday day and month after cleaning
median_day <- median(C1survey_csv$bday, na.rm = TRUE)
median_month <- median(C1survey_csv$bmonth, na.rm = TRUE)
median_day
median_month
```

```{r, chunk17}
# 6. a. Create new variable bseason based on meteorological seasons

C1survey_csv <- C1survey_csv %>%
  mutate(
    bseason = if_else(bmonth %in% c(12, 1, 2), "Winter",
               if_else(bmonth %in% 3:5, "Spring",
               if_else(bmonth %in% 6:8, "Summer",
               if_else(bmonth %in% 9:11, "Fall", NA_character_)))))

```
```{r, chunk18}
# 6.b. Check the coding by creating a table of bmonth vs bseason

table(C1survey_csv$bmonth, C1survey_csv$bseason, useNA = "always")

```
```{r, chunk19}
#Check coding: table of months (rows) vs seasons (columns)
season_table <- table(C1survey_csv$bmonth, C1survey_csv$bseason, useNA = "always")
season_table
```
```{r, chunk20}
# 6.c. Sum the number of classmates in each season using addmargins
addmargins(season_table, 2)  # sum across columns


```
```{r, chunk21}

# 7. Choose one or more variables for own analysis
# I choose like cats and like dogs variables
# Hypothesis: Most people who like cats also like dogs.

# Subset the dataset to only people who like cats
like_cats_subset <- C1survey_csv %>%
  filter(like_cats == "Yes")

# Calculate the proportion of these people who also like dogs
mean(like_cats_subset$like_dogs == "Yes", na.rm = TRUE)

```


